\documentclass{article}
\usepackage[utf8]{inputenc}
\usepackage{amsmath,amsthm,amssymb,tikz-cd}
\usepackage[T1]{fontenc}
\usepackage{enumitem}
\usepackage{fullpage}
\usepackage{setspace}
\usepackage{mathtools}
\usepackage[bookmarks]{hyperref}
\newcommand{\R}{\mathbb{R}}
\newcommand{\Z}{\mathbb{Z}}
\newcommand{\C}{\mathbb{C}}
\newcommand{\F}{\mathbb{F}}
\newcommand{\aut}{\operatorname{Aut}}
\newcommand{\inn}{\operatorname{Inn}}
\newcommand{\ga}{\langle a \rangle}
\renewcommand{\phi}{\varphi}

\onehalfspacing
\setlength{\parskip}{0em}
\setlist{listparindent=\parindent,parsep=0pt}

\begin{document}

\begin{center}
    {\Large MATH 113 Summary of Facts} \vspace{0.5em}
    \\ Cheng, Feng
    \vspace{-0.5em}
\end{center}

\section{Groups and Subgroups}
\begin{itemize}
    \item Left and right cancellation laws in groups
    \item Unique solution to all linear equations, and thus every element appears exactly once in every row and column of a group table
    \item Unique identity and inverse
    \item $(ab)^{-1}= b^{-1}a^{-1}$
    \item There are one group of order 1/2/3, up to isomorphism. There are two groups of order 4, up to isomorphism, $\Z_4$ and Klein 4-group $V$. Groups of order less than 6 are all abelian. The smallest nonabelian group is of order 6, and $S_3$ is a common counterexample to many claims.
    \item The only idempotent element of a group is the identity.
    \item $\Z_4$ has one nontrivial proper subgroup $\{0,2\}$, while $V$ has three: $\{e,a\}, \{e,b\},\{e,c\}$.
    \item To check if $H \subseteq G$ is a subgroup, we check $H$ is closed, $H$ has an identity, and the inverse exists for every element of $H$. Associativity is inherited from the whole group.
    \item $H = \{a^n \mid n \in \Z\}$ is the smallest subgroup of $G$ containing $a \in G$, called the cyclic subgroup of $G$ generated by $a$. $H$ can be denoted by $\langle a \rangle$. If $G$ is the whole group, then $a$ generates $G$ and $G$ is a cyclic group.
    \item Every cyclic group is abelian.
    \item A subgroup of a cyclic group is cyclic by the division algorithm. This implies that the subgroups of $\Z$ are the groups $n\Z$ (they are all the cyclic subgroups of $\Z$ generated by elements of $\Z$).
    \item The positive generator of $H = \{ar+bs \mid a,b \in \Z\}$ under addition is $\gcd(r,s)$.
    \item If $G = \langle a \rangle$ is infinite, then it is isomorphic to $\Z$ under addition; if it is finite, then it is isomorphic to $\Z_n$ under $+_n$. In the proof we show that for finite $\langle a \rangle$, the smallest number $m$ such that $a^m = e$ is its order. For infinite $\langle a \rangle$, m does not exist.
    \item $G = \langle a \rangle$ of order $n$, then $a^s$ generates a cyclic subgroup of order $\frac{n}{\gcd(n,s)}$. Also $\langle a^s \rangle = \langle a^t \rangle$ iff $\gcd(s,n)=\gcd(t,n)$.
    \item It follows that $a$ generates $G$ of order $n$, then all other generators of $G$ are $a^r$, where $r$ and $n$ are coprime.
    \item Arbitrary intersection of subgroups $H_i$ of a group $G$ is a subgroup of $G$.
    \item The smallest subgroup of G containing $\{a_i \mid i \in I\}$ is the subgroup generated by this generating set. It has the elements of $G$ that are the product of integral powers of the $a_i$'s, and one $a_i$ may appear several times in the product because $G$ is not necessarily abelian.
\end{itemize}

\section{Permutation, Cosets, and Direct Products}
\begin{itemize}
    \item A permutation of a set $A$ is a bijective function $\phi: A \to A$.
    \item The set of permutation $S_A$ is a group under permutation multiplication (composition).
    \item Permutation $S_n$ on $\{1,2,\dots,n\}$ is isomorphic to all permutations of order $n$, and can thus be taken as the prototype for permutations of order $n$.
    \item $S_3 \simeq D_3$ under renaming
    \item For one-to-one function $\phi: G \to G'$ satisfying the homomorphism property, the image $\phi(G)$ is a subgroup of $G'$ and $\phi(G) \simeq G$. (This is actually true for algebraic structures in general because isomorphism is renaming.) We use this to show Cayley's theorem that every group is isomorphic to a group of permutations.
    \item Given $\sigma \in S_A$, the equivalence relation $\sim_\sigma$ given by $a \sim b$ ($a,b \in A$) if $b = \sigma^n(a)$ for some $n \in \Z$. This particular equivalence relation on the set $A$ determined by permutations on $A$ partitions the set $A$ into equivalence classes. The equivalence class determined by $\sim_\sigma$ is called the \textit{orbits} of $\sigma$.
    \item We now restrict to permutation on a finite set $A$, and we may take $S_n$ as the representative. These equivalence classes are exactly the \textit{cycles} of a permutation. Since $A$ is a finite set, it is impossible that no element $a \in A$ will never return to itself. Thus, an equivalence class of $A$ is a circle that starts from an element $a \in A$ and returns to $a$.
    \item Thus, every permutation on a finite set can always be broken down into a product of disjoint \textit{cycle permutations} (which only switches the ordering of elements on the orbit by 1).
    \item Every cycle can be broken down into transpositions: $$(a_1, a_2, \dots, a_n) = (a_1, a_n) (a_1, a_{n-1}) \dots (a_1,a_2).$$ Therefore, any permutation of a finite set with at least elements is a product of transpositions.
    \item If a permutation can be written into a product of even (odd, resp.) number of transpositions, then it can only be written into a product of even (odd, resp.) number of transpositions. We call them \textit{even permutations} (and \textit{odd permutations}), denoted by $A_n$ (and $B_n$).
    \item One can construct a bijective map between $A_n$ and $B_n$ (for $n \geq 2$). Therefore, $|A_n| = |B_n| = (n!)/2$. It can be easily verified that $A_n$ is a subgroup of $S_n$, and we call it the \textit{alternating group}.
    \item For $G$ and $H \leq G$, we can define a relation $\sim_L$ on $G$ by $$a \sim_L b \quad \text{if} \quad a^{-1}b \in H,$$ and another relation $\sim_R$ on $G$ by $$a \sim_L b \quad \text{if} \quad ab^{-1} \in H.$$ These two are equivalence relations, and divide $G$ into \textit{left} and \textit{right cosets}, respectively. Left and right cosets are each of the form $aH$ and $Ha$. If $aH = Ha$ for all $a \in G$, we call $H$ a \textit{normal subgroup} of $G$.
    \item We focus mainly on left cosets. A coset $aH$ can have multiple representatives as long as the other representative $b$ is in $aH$ (or equivalently, $a^{-1}b \in H$).
    \item Lagrange's Theorem states that the order of a subgroup $H$ of a finite group $G$ is the divisor of the order of the whole group. This is because the cardinalities of the cosets and the order of the subgroup are the same. We prove this by the natural bijective map $\phi(h) = gh$ from $H$ to $gH$.
    \item A prime-order group must be cyclic. If one considers the cyclic subgroup generated by $a \not= e$, then $|\ga| \geq 2$ and thus must be $p$.
    \item The number of left cosets of $H$ in $G$ is the \textit{index} $(G:H)$ of $H$ in $G$. The product rule $(G:K) = (G:H)(H:K)$ holds for a chain of subgroup, regardless of whether the groups involved are finite or not.
    \item The entry-wise direct product $\prod_{i} G_i$ of groups $G_i$ forms a new group. Rearrangement of factors yields the same group of direct product, up to isomorphism.
    \item $\Z_m \times \Z_n$ is cyclic and thus isomorphic to $\Z_{mn}$ if and only if $\gcd(m,n) = 1$. Obviously this can be extended to direct products of higher dimensions.
    \item For $(a_1, \dots, a_n) \in \prod_{i=1}^n G_i$, the order of this element is the lcm of the orders of each individual $a_i$'s.
    \item The fundamental theorem of finitely generated abelian groups states that every finitely generated abelian groups $G$ is isomorphic to a direct product of cyclic groups in the form $$\Z_{(p_1)^{r_1}} \times \Z_{(p_2)^{r_2}} \times \cdots \times \Z_{(p_n)^{r_n}} \times \Z \times \Z \times \cdots \times \Z,$$ where the $\Z$'s at the end have a fixed number called the \textit{Betti number} of $G$, and the $p_j$'s at the front do not need to be distinct.
    \item A group $G$ is \textit{decomposable} if it is isomorphic to a direct product of two nontrivial subgroups. The finite indecomposable abelian groups are exactly the ones with order a power of a prime.
    \item If $m \mid |G|$, where $G$ is a \textbf{finite abelian group}, then $G$ has a subgroup of order $m$. Consider $$\langle (p_1)^{r_1 - s_1} \rangle \times \langle (p_2)^{r_2 - s_2} \rangle \times \cdots \times \langle (p_n)^{r_n - s_n} \rangle.$$
    \item If $m$ is a square-free integer, then abelian group of order $m$ is cyclic.
\end{itemize}

\section{Homomorphisms and Factor Groups}
\begin{itemize}
    \item A surjective group homomorphism $\phi$ from an abelian group $G$ to another group $G'$ tells us $G'$ is abelian. (This is the preservation of algebraic structure.)
    \item It suffices to show from the homomorphism property that $\phi(e_G) = e_{G'}$, $\phi(a^{-1}) = (\phi(a))^{-1}$.
    \item The image of a subgroup of $G$ is a subgroup of $G'$. The preimage of a subgroup of $G'$ is a subgroup of $G$. It follows directly that the kernel and image of a group homomorphism is the respective subgroups of $G$ and $G'$. Furthermore, $\ker \phi$ is a normal subgroup of $G$.
    \item $\phi$ is one-to-one if and only if its kernel only contains $e$. Thus, to prove isomorphism, one checks the homomorphism property, the kernel, and onto.
    \item There are several equivalent conditions to check that a subgroup is normal. Also, every subgroup of an abelian group is normal.
    \item The coset multiplication operation on $H \leq G$ is well-defined if and only if $H \trianglelefteq G$.
    \item Given $\phi: G \to H$ and the natural surjective homomorphism $\pi: G \to G/N$ with kernel $N$, $\tilde{\phi}: G/N \to H$ defined by $$\tilde{\phi}(gN) = \phi(g)$$ is well-defined if and only if $N \leq \ker \phi$.
    \item The full version of the fundamental theorem of group homomorphism states that, given a group homomorphism $\phi: G \to H$, let $N$ be a normal subgroup in $G$ and $\pi: G \to G/N$ be the natural surjective homomorphism. If $N \leq \ker \phi$, then there exists a unique homomorphism $\tilde{\phi}: G/N \to H$ such that $\phi = \tilde{\phi} \circ \pi$.
    \begin{itemize}
        \item In particular, if we let $\ker \pi = N = \ker \phi $, then $\ker \tilde{\phi}$ contains only the identity and is thus injective. Therefore, $\tilde{\phi}: G/N \to \phi(G)$ is an isomorphism.
    \end{itemize}
    \item The automorphism group $\aut(G) \coloneqq \{\phi: G \to G \mid \phi \text{ is an isomorphism}\}$. In particular $i_g: G \to G$ given by $i_g(x) = g x g^{-1}$ is an automorphism of $G$, called the inner automorphism of $G$. These $i_g$'s form $\inn(G) \coloneqq \{\phi \in \aut(G) \mid \phi = i_g \text{ for some } g \in G\}$, which turns out to be a normal subgroup of $\aut(G)$.
    \item $\operatorname{Out}(G) \coloneqq \aut(G)/\inn(G)$.
    \item $\phi: G \to \inn(G)$ is a surjective group homomorphism with kernel $Z(G) \coloneqq \{g \in G \mid g x g^{-1} = x \text{ for all } x \in G\}$, called the \textit{center} of $G$. Hence $G/Z(G) \simeq \inn(G)$.
    \item A group is \textit{simple} if it has no proper nontrivial normal subgroup.
    \item $A_n$ ($n \geq 5$) is simple.
\end{itemize}

\section{Rings and Fields}
\begin{itemize}
    \item The definition of rings given in the book does not assume the existence of the multiplicative identity. A ring $\langle R,+,\cdot \rangle$ is a set $R$ endowed with two binary operations addition and multiplication, such that:
    \begin{enumerate}%[label=(\alph*)]
        \item $\langle R, + \rangle$ is an abelian group;
        \item $R$ is associative under multiplication;
        \item Left and right distributive laws hold.
    \end{enumerate}
    \item A \textit{0 divisor} of 
    \item A polynomial $f(x)$ 
\end{itemize}
\end{document}